Supplementary Information

### 2020-09-14 doc/manuscript/MEE_manuscript.Rmd
started incorporating JLB feedback

[Kristina J. Anderson-Teixeira^1,2^*](0000-0001-8461-9713),
Valentine Herrmann^1^,
Christy Rollinson^3^,
Bianca Gonzalez^1^,
Erika B. Gonzalez-Akre^1^,
Neil Pederson^4^,
Camille Piponiot^1,2^,
Ross Alexander,
Craig D. Allen,
Raquel Alfaro-Sánchez,
Tala Awada,
Jennifer Baltzar,
Patrick Baker,
Sarayudh Bunyavejchewin,
Paolo Cherubini,
Justin Cooper,
Stuart Davies^2^,
Ryan Helcoski^1^,
[Jakub Kašpar^cz^](0000-0003-1780-6310),
Kamil Kral,
James Lutz,
Ellis Q. Margolis,
Justin Maxwell^jm^,
Sean McMahon^2^,
Helene Muller-Landau^2^,
Sabrina Russo,
[Pavel Šamonil^cz^](0000-0002-7722-8797),
Anastasia Sniderhan,
Alan J. Tepley^1,at^,
[Ivana Vašíčková^cz^](0000-0002-6070-5956), 
Mart Vlam,
Peter Zuidema

@@

@@

**Tree size (most commonly diameter breast height, $DBH$) is among the most important variables affecting tree growth rate [e.g., @muller-landau_testing_2006; @foster_predicting_2016] and its climate sensitivity [e.g., @bennett_larger_2015; @mcgregor_tree_nodate], yet traditional methods do not characterize its effect or its potential interactions with climate.** 
For dendrological studies aimed at deciphering climate signals, $DBH$ is not typically a variable of interest, and its influence is removed through detrending (DENDRO_REFS).
While convenient for identifying climate signals [(DENDRO_REFS)], this approach is not optimal for subsequent inference of the climate sensitivity of forest productivity.
Although climate correlations can be transformed to climate sensitivity [*sensu* @charney_observed_2016] and scaled to characterize the climate sensitivity of $ANPP_{woody}$ based on the scaling of $\Delta r$ with $DBH$ [@helcoski_growing_2019], they cannot be used to characterize known interactive effects of $DBH$ and climate on tree growth.
For example, larger trees tend to exhibit greater growth reductions in drought years [@bennett_larger_2015; @mcgregor_tree_nodate], and smaller trees can be more sensitive to temperature (*Rollinson et al. in review*).
However, the removal of $DBH$ signals via detrending makes it impossible to account for such size differences in climate sensitivity. 
To use tree-rings to predict tree growth and $ANPP_{woody}$, we need models that include $DBH$ and its potential climate interactions.

>>>>

[Kristina J. Anderson-Teixeira^1,2^*](0000-0001-8461-9713),
Valentine Herrmann^1^,
Christy Rollinson^3^,
Bianca Gonzalez^1^,
Erika B. Gonzalez-Akre^1^,
Neil Pederson^4^,
Camille Piponiot^1,2^,
Ross Alexander,
Craig D. Allen,
Raquel Alfaro-Sánchez^wl^,
Tala Awada,
Jennifer Baltzer^wl^,
Patrick Baker,
Sarayudh Bunyavejchewin,
Paolo Cherubini,
Justin Cooper,
Stuart Davies^2^,
Ryan Helcoski^1^,
[Jakub Kašpar^cz^](0000-0003-1780-6310),
Kamil Kral,
James Lutz,
Ellis Q. Margolis,
Justin Maxwell^jm^,
Sean McMahon^2^,
Helene Muller-Landau^2^,
Sabrina Russo,
[Pavel Šamonil^cz^](0000-0002-7722-8797),
Anastasia Sniderhan^wl^,
Alan J. Tepley^1,at^,
[Ivana Vašíčková^cz^](0000-0002-6070-5956), 
Mart Vlam,
Peter Zuidema

@@

wl. Biology Department, Wilfrid Laurier University, 75 University Ave W, Waterloo, ON, N2L 3C5

@@

**Tree size (most commonly diameter breast height, $DBH$) is among the most important variables affecting tree growth rate [e.g., @muller-landau_testing_2006; @foster_predicting_2016] and its climate sensitivity [e.g., @bennett_larger_2015; @mcgregor_tree_nodate], yet traditional methods do not characterize its effect or its potential interactions with climate.** 
For dendrological studies aimed at deciphering climate signals, $DBH$ is not typically a variable of interest, and its influence is removed through detrending (DENDRO_REFS).
Moreover, many studies constrain sampling to only larger size classes.
While convenient for identifying climate signals [(DENDRO_REFS)], this approach is not optimal for subsequent inference of the climate sensitivity of forest productivity.
Although climate correlations can be transformed to climate sensitivity [*sensu* @charney_observed_2016] and scaled to characterize the climate sensitivity of $ANPP_{woody}$ based on the scaling of $\Delta r$ with $DBH$ [@helcoski_growing_2019], they cannot be used to characterize known interactive effects of $DBH$ and climate on tree growth.
For example, larger trees tend to exhibit greater growth reductions in drought years [@bennett_larger_2015; @mcgregor_tree_nodate], and smaller trees can be more sensitive to temperature (*Rollinson et al. in review*).
However, the removal of $DBH$ signals via detrending makes it impossible to account for such size differences in climate sensitivity. 
To use tree-rings to predict tree growth, $ANPP_{woody}$, and forest dynamics, we need models that include $DBH$ and its potential climate interactions.
